---
title: "p8105_hw3_alc2279"
author: "Annie Clark"
date: "October 6, 2018"
output: github_document
---

# Problem 0
This “problem” focuses on structure of your submission, especially the use git 
and GitHub for reproducibility, R Projects to organize your work, R Markdown 
to write reproducible reports, relative paths to load data from local files, 
and reasonable naming structures for your files.

To that end:

* create a public GitHub repo + local R Project; we suggest naming this repo / 
directory p8105_hw3_YOURUNI (e.g.  p8105_hw3_ajg2202 for Jeff), but that’s not 
required
* create a single .Rmd file named p8105_hw3_YOURUNI.Rmd that renders to 
github_document

Your solutions to Problems 1, 2, and 3 should be implemented in your .Rmd file, 
and your git commit history should reflect the process you used to solve these 
Problems.

For this Problem, we will assess adherence to the instructions above regarding 
repo structure, git commit history, and whether we are able to knit your .Rmd 
to ensure that your work is reproducible. Adherence to appropriate styling and 
clarity of code will be assessed in Problems 1+ using the homework style rubric.

This homework includes figures; the readability of your embedded plots (e.g. 
font sizes, axis labels, titles) will be assessed in Problems 1+.

```{r}
library(tidyverse)

library(p8105.datasets)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

# Problem 1

## Part 1: Load and tidy the data
This problem uses the BRFSS data. DO NOT include this dataset in your local data
directory; instead, load the data from the  p8105.datasets package.

First, do some data cleaning:

* format the data to use appropriate variable names;
* focus on the “Overall Health” topic
* include only responses from “Excellent” to “Poor”
* organize responses as a factor taking levels from “Excellent” to “Poor”

```{r}
brfss_data = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(state = locationabbr, state_county = locationdesc, proportion = data_value) %>% 
  filter(topic == "Overall Health", 
         response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>% 
  select(-(class:question), -(confidence_limit_low:geo_location)) %>% 
  mutate(response = forcats::fct_relevel(response)) %>% 
  arrange(response)

##don't actually need to filter by responses because when you group by, you see that the only response for overall health are excellent through poor, but keeping this in case the data changes and it is still a requirement by collaborators

  
##Question - only include excellent through poor? those are the only responses?
##Question - does 'organize as factor...' mean to recode as a factor and then arrange in ascending order?
```

## Part 2: Answer some questions

Using this dataset, do or answer the following (commenting on the results of 
each):

* In 2002, which states were observed at 7 locations?

```{r}
brfss_data %>% 
  filter(year == "2002") %>% 
  group_by(state) %>% 
  summarize(n_locations = n_distinct(state_county)) %>% 
  filter(n_locations == 7) %>% 
  pull(state)
```

* Make a “spaghetti plot” that shows the number of observations in each state 
from 2002 to 2010.

```{r}
##brfss_data %>% 
  ##group_by(state) %>% 
  ##summarize(n_observations = n()) %>% 
  ggplot(brfss_data, aes(x = year, y = sample_size, color = state)) + 
    geom_line()

##Question - wut. is going on with this graph.
```

* Make a table showing, for the years 2002, 2006, and 2010, the mean and standard 
deviation of the proportion of “Excellent” responses across locations in NY State.

```{r}
brfss_data %>% 
  group_by(state_county) %>% 
  filter(year %in% c("2002", "2006", "2010"),
         state == "NY",
         response == "Excellent") %>%
  summarize(excellent_mean = mean(proportion, na.rm = TRUE), 
            excellent_sd = sd(proportion, na.rm = TRUE))

##Question - group by location and year or just location?
##pulling back NaN because there is only one point estimate; cannot calculate sd?
```

* For each year and state, compute the average proportion in each response category 
(taking the average across locations in a state). Make a five-panel plot that shows, 
for each response category separately, the distribution of these state-level 
averages over time.

```{r}
brfss_data %>% 
  group_by(year, state, response) %>% 
  summarize(response_mean = mean(proportion, na.rm = TRUE)) %>% 
  ##spread(key = response, value = response_mean) %>% 
  ggplot(aes(x = response, y = response_mean)) + 
    geom_boxplot()

##Question - clarify what they're actually asking? did i do it?
```

# Problem 2

## Part 1: Load the data
This problem uses the Instacart data. DO NOT include this dataset in your local 
data directory; instead, load the data from the p8105.datasets package (it’s 
called instacart).

```{r}
instacart_data = instacart
```

## Part 2: EDA - Describe the data

### Prompt

The goal is to do some exploration of this dataset. To that end, write a short 
description of the dataset, noting the size and structure of the data, describing 
some key variables, and giving illstrative examples of observations. 

### Response


## Part 3: EDA - Some questions

Then, do or answer the following (commenting on the results of each):

* How many aisles are there, and which aisles are the most items ordered from?

```{r}

```

* Make a plot that shows the number of items ordered in each aisle. Order aisles 
sensibly, and organize your plot so others can read it.

```{r}

```

* Make a table showing the most popular item aisles “baking ingredients”, “dog 
food care”, and “packaged vegetables fruits”

```{r}

```

* Make a table showing the mean hour of the day at which Pink Lady Apples and 
Coffee Ice Cream are ordered on each day of the week; format this table for 
human readers (i.e. produce a 2 x 7 table).

```{r}

```

# Problem 3

## Part 1: Load the data
This problem uses the NY NOAA data. DO NOT include this dataset in your local 
data directory; instead, load the data from the p8105.datasets package (it’s 
called ny_noaa).

```{r}
noaa_data = ny_noaa
```

## Part 2: EDA - Describe the data

### Prompt

The goal is to do some exploration of this dataset. To that end, write a short 
description of the dataset, noting the size and structure of the data, describing 
some key variables, and indicating the extent to which missing data is an issue. 

### Response

## Part 3: EDA - Answer some questions

Then, do or answer the following (commenting on the results of each):

* Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable 
units. For snowfall, what are the most commonly observed values? Why?

```{r}

```

* Make a two-panel plot showing the average temperature in January and in July in 
each station across years. Is there any observable / interpretable structure? 
Any outliers?

```{r}

```

* Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that 
a scatterplot may not be the best option); and (ii) make a plot showing the 
distribution of snowfall values greater than 0 and less than 100 separately by 
year.

```{r}

```

